Collaborative Learning with Dirichlet Process Clustering for Rapid Online Adaptation in Robotics

### 2023-09-05 paper/methodology.tex
Sep 5

\documentclass[journal]{IEEEtran}
%\documentclass{article}
%\usepackage[margin=1in]{geometry}
%\usepackage{multicol}

\title{Collaborative Learning with Dirichlet Process Clustering for Rapid Online Adaptation in Robotics}
\author{Runjun Mao \\ August 12, 2023 \\ Supervised by Antoine Cully}
%\date{August 12, 2023}

\usepackage{bm}
%\usepackage{amssymb}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\renewcommand{\algorithmiccomment}[1]{\hfill $\triangleright$ #1}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\hypersetup{hidelinks}
%\usepackage[colorlinks, linkcolor=blue, citecolor=green, urlcolor=red]{hyperref}

\begin{document}
\maketitle

\begin{abstract}
\noindent
Robots in deployment can face unforeseeable distortion in their motion due to unseen environment or hardware failures. 
Model-based online adaptation learns this distortion from the interactions with the current environment and corrects its motions accordingly. However, this can be very challenging as it is hard to select the most suitable model and its hyper-parameters for the current dynamics with little prior knowledge.
If we could leverage the plentiful historical real-world interactions, we may build statistical models for the common types of dynamics faced by the robot, hence providing guidance for adaptations. 
This paper developed a method to enable fast online adaptation for the robot by grouping up historical  interaction data into clusters. 
The historical data can be collected from several robots of the same design performing different tasks in different environments, hence allowing learning in a collaborative manner. 
The method takes an non-parametric approach of novelty by modelling the data distribution with infinite mixture of Gaussian Processes and conducting clustering using Dirichlet Processes. 
This achieves unbiased training while ensures high efficiency in terms of both data and computation during deployment. 
The experiment is conducted on a simulated robot to compare with the previous work for online adaptation. 
The method accurately groups up the collaboratively collected data without any knowledge of their sources, and it offers a significant improvement over the baseline in both the efficiency of learning the distortion and the capability of adaptation.

\end{abstract}

%\begin{multicols}{2}

\section{Introduction}
Locomotion controls of legged robots are typically too complex to be designed manually.
In recent years, deep Reinforcement-Learning (deep RL) has made huge success in game playing
\cite{alphaGo, alphaStar} and is regarded promising for application in robotics. However, training such well performing agents requires enormous amounts of interactions with the environment, meaning that this can only 
be done in simulation where we can model the interaction and collect the training data much faster than in real-time. This raises a problem for application in robotics since there is always a gap between simulation and reality, and the policy trained in simulation could suffer from this sim-to-real transfer. Another problem is that any unforeseen changes to the environment dynamics could lead to severely degraded performance. These deep RL trained agents require extra fine-tuning or even retaining to retain their capabilities.
For example, the team of OpenAI Five \cite{openAI5} had to conduct three additional amendments in the architecture followed by extra trainings to incorporate small version updates in Dota 2 during the development of the agent.
For game-playing agents, these drawbacks are tolerable as there is no sim-to-real problem, and the environment is overall stationary. For robotics however, sim-to-real must be considered, and the environment dynamics is typically non-stationary as the robot can be travelling between different terrain, carrying different payloads or suffering from damages in its hardware.

It would be better if we may learn the true dynamics from the interactions with the current dynamics and
correct our policy accordingly. This is a model-based reinforcement learning (MBRL) solution that allows robots to learn policies with lesser interactions by learning a dynamics model and use it optimize the policy \cite{MBRL, black_box_search, policy_search}. 
However, the amount of data required to learn a model typically scales exponentially with the dimensionality of the input space \cite{curse_of_dim}, hence it is not suitable for online adaptation. 
The promising approach to address this is to use a repertoire-based method, where we learn a repertoire of elementary policies (e.g., one policy for turning left, one for moving forward, etc.) using quality diversity optimisation \cite{QD} to fully cover the task space (e.g., the 2D displacement we want the robot to make). 
The repertoire-based method is a hierarchical control method that keeps a variety of elementary policies and treats each policy as an action. 
We then give each elementary policy a behaviour descriptor (BD) to quantify the way it acts.
This enables us to skip the low-level states and actions (i.e., joint encoders and torques for all motors) and to work with the high-level behaviour space that has much smaller dimensionality.
During the adaptation, we learn the distortion (called transformation model) in the behaviour space from real-world interactions. Thanks to the reduced dimensionality, the transformation model can be learned in real-time.
Instead of optimising the policy with this learned model, we use a repertoire-based control that first predicts the outcomes for each elementary policy in the repertoire and then simply chooses the one with the predicted outcome that is most aligned with the current goal. 
This is made available for keeping a repertoire of policies rather than a single policy, so that the robot can adapt to the distortion by finding the corresponding policies that compensate the distortion (for example the robot aims to go forward and distortion is left, then a policy originally aiming right-front can be used for going forward instead).

The repertoire-based method has achieved very impressing results in task-solving \cite{EvoRBC} and online damage recovery \cite{RTE}. 
But this method heavily relies on having a good model that is both data efficient and suitable for modelling the distortion. 
The most commonly used model is Gaussian Process (GP) \cite{GP}. The prior mean function and the kernel are the most important factors of GP. 
If no prior knowledge is given about the dynamics, these  have to be selected manually according to experience.
If the distortion is very large (like broken legs) or complex, the GPs can be very inefficient or even misleading.
If we have data of real-world interactions across several different environments, we can build a GP for each of the situations and then teach the robot to identify the one that mostly explains its current situation. However, such real-world data are expensive to collect and could lead to covariate shift (the collected data distribution is different from that during deployment).
A promising solution that is to leverage the historical data of real-world interactions collected during deployment. 
After performing each task, the interaction data can be uploaded to an archive system where they will be analysed. 
Thus, the data doesn't have to be collected on purpose, and the acquired data distribution is completely unbiased. 
This also allows the robots to learn to adapt in a collaborative manner.
If we have multiple robots performing different tasks in different environments, we can quickly get an overview of the  common types of dynamics and build the adaptation strategy accordingly. 
If a robot encounters a new environment, with the interaction data uploaded, all the robots will be able to quickly adapt to this environment.
Finally, considering the scalability of such system, we can group the data into clusters based on their dynamics and assign a GP (with the corresponding mean and kernel) to each cluster. This prevents the adaptation strategy from becoming computationally expensive after we have collected data from thousands of tasks.

There are many challenges needs to be addressed to enable this collaborative learning. First, traditional clustering algorithms like k-nearest neighbours \cite{knn} and k-means \cite{k-means} need to specify the number of clusters. While in our case, this number is not even fixed as we might observe more clusters as we collect more data.
Second, we know that each task corresponds to a single environment, and hence we are essentially clustering the tasks instead of clustering the data points.
However, since some tasks are tougher than others, different tasks generate different amount of data. 
As a result, we will be clustering data of inconsistent sizes.
Third, since there is a large repertoire of policies, only a small subset of them will be executed while performing each task. Hence, it is very likely that the data collected from two tasks have no overlapping policies, making it hard to determine whether they have similar dynamics and whether they should be grouped together.
Moreover, after we have found the clusters, the data in each cluster can never cover the entire repertoire. Hence it is difficult to determine the prior mean for the policies that are not present in this cluster.
This paper aims to solve all the above problems by taking an non-parametric clustering method and using a uniquely designed prior mean function.

\section{Related Works}
\subsection{Quality-Diversity Optimization}
The repertoire-based method uses quality diversity (QD) optimization to generate a repertoire of diversified and high-quality locomotion behaviours. 
The most popular method for doing QD is called MAP-Elites \cite{QD, Map-Elites}, where we tabularize the task-space (for example, discretize the displacement of the robot along the x, y axis) into grids. 
Each grid corresponds to a policy that leads the robot to this part of task space. 
Initially the grids are empty, as we have no policies at start, and will be filled during a process of stochastic optimization inspired by natural evolutionary. 
We start from a few random policies, which are typically neural networks, and execute them in the simulator to get their outcomes in the task space. 
We also assign a performance score (called fitness) to each policy to quantify the efficiency of the execution. Typically, the fitness can be based on the energy consumption \cite{Q-Dax} or the shape of the trajectory \cite{evorbc_conf}.
Each of these outcomes is mapped to one of the girds (hence these grids are filled). 
We then begin our main loop of policy generation. 
In each iteration we select a few policies corresponding to the filled grids as the parents. We then conduct crossover on the parents to mutate their genotypes (for neural networks, these are just the weights of the neurons) to get their off-springs. 
The crossover operation should balance the exploitation and exploration by ensuring some inheritance of the parents to preserve good genes as well as some random variations for promoting diversity.
We then evaluate the off-springs in the simulator to get their outcomes in the task space and identify the grids they belong to.
If a grid is already occupied by another policy, we only keep the one with the higher fitness; if it is empty, we fill this grid with the corresponding off-spring. 
The pseudo code for MAP-Elites QD algorithm is presented below.

\begin{algorithm}
\caption{MAP-Elites}
\begin{algorithmic}
\STATE \textbf{procedure} MAP-Elites
\STATE Discretize the task space into grids
\STATE $\mathcal{R} \leftarrow \emptyset$ \COMMENT{Initialize repertoire}
\FOR{$\bm{i} = 1 \rightarrow G$}
\STATE $\theta \leftarrow $ generate{\_}random{\_}policy()
\STATE $\Theta$, $f$ $\leftarrow$ evaluate{\_}policy($\theta$) \COMMENT{Get outcome and fitness}
\STATE add{\_}to{\_}repertoire($\Theta$, $f$, $\theta$, $\mathcal{R}$) \COMMENT{Update repertoire}
\ENDFOR
\FOR{$\bm{i} = 1 \rightarrow T$}
\STATE $\vartheta \leftarrow$ select{\_}parents($\mathcal{R}$)
\STATE $\bm{\theta} \leftarrow$ crossover($\vartheta$, $N$) \COMMENT{Generate $N$ off-springs}
\FOR{$\bm{j} = 1 \rightarrow N$}
\STATE $\Theta$, $f$ $\leftarrow$ evaluate{\_}policy($\bm{\theta_j}$)
\COMMENT{Off-spring evaluation}
\STATE add{\_}to{\_}repertoire($\Theta$, $f$, $\bm{\theta_j}$, $\mathcal{R}$)
\ENDFOR
\ENDFOR
\end{algorithmic}
\label{MAP-Elites}
\end{algorithm}
\noindent
Evaluating the policies in the simulation is the major source of computationally expense for the MAP-Elites.
A merit of this algorithm is that the inner for loop of off-spring evaluation can be conducted in parallel, and the repertoire update can be made after having collected all the results of the off-springs.
This enables us to generate a large number of off-springs in each iteration and evaluate them in parallel, fully leveraging the power of modern Cluster computing \cite{Q-Dax}.
By running the algorithm for a few thousand iterations, we will end up having a repertoire of high-performing policies well-covering the task space.

\subsection{Gaussian Process}
The Gaussian Process (GP) is a powerful non-parametric machine learning algorithm. 
It assumes the prior distribution of the outcomes to be a multivariate normal distribution with a prior mean $\mu$ and a covariance matrix $\Sigma$. 
\begin{equation}
f(\bm{y}) = \frac{exp(
-\frac{1}{2}(\bm{y} - \bm{\mu})^T \bm{\Sigma^{-1}} (\bm{y} - \bm{\mu})
)}
{
\sqrt{(2\pi)^N \det(\bm{\Sigma})}
}
\label{multi_normal}
\end{equation}
In contrast to regular multivariate normal distribution, GP allows the prior mean and the covariance matrix to be input dependent. The
prior mean becomes a function $\mu(\bm{x})$, and the covariance between two data points is given by a covariance function 
$k(\bm{x_1}, \bm{x_2})$ called kernel. 
Following the fact that data points closer to each other are more correlated, this covariance starts from the prior variance and asymptotically decreases to zero as the distance between the two points tends to infinity. 
To account for the fact the observations might be noisy, we can incorporate the noise into the covariance matrix by adding the variance of the noise on the diagonal.
\begin{equation}
\bm{\Sigma} = \bm{K}(\bm{x_{1:N}}, \bm{x_{1:N}}) + \sigma_n^2 \mathbf{I}
\label{covariance_matrix}
\end{equation}

This allows us to build reasonable joint prior distribution for the values of any well-behaved functions.
To make prediction with GP, we first use the prior mean function and kernel to build up the joint distribution as the prior. 
Then we can calculate the posterior for the data we wish to evaluate conditioned on the ones we have observed using Bayes rule.
Since the Gaussian distribution is a conjugate distribution, the posterior is still a Gaussian with mean and variance given by \cite{GP, GP_posterior}:
\begin{equation}
\begin{gathered}
\mu_*(x) = \mu(x) + 
\bm{\Sigma}_{N, x}^T
\bm{\Sigma}_{N, N}^{-1}
(\bm{y}_{1:N} - \bm{\mu}(\bm{x}_{1:N}))
\\
\sigma^2_*(x) = \sigma^2(x) -  
\bm{\Sigma}_{N, x}^T
\bm{\Sigma}_{N, N}^{-1}
\bm{\Sigma}_{N, x}
\end{gathered}
\label{GP_posterior}
\end{equation}
where $\bm{\Sigma}_{N, N}$ denotes the covariance matrix of the observed data, and $\bm{\Sigma}_{N, x}$ denotes the column vector with entry on each row equal to the covariance between 
each observed data and the data we hope to evaluate.

Typically, the prior mean is just a fixed constant like zero, representing the prior knowledge we have about the outcome when there is no neighbouring data to refer to.
The typical choice of kernel can be RBF kernel that models the correlations to decrease in the form of a Gaussian function.
\begin{equation}
k(\bm{x_i}, \bm{x_j}) = \alpha_0 e^{-\frac{1}{2} d_{ij}^2}
\label{RBF_kernel}
\end{equation}
RBF kernel assumes the function is infinitely differentiable, which may be factually incorrect. Hence, it might be more desirable to use the Matern kernel:
\begin{equation}
k(\bm{x_i}, \bm{x_j}) = \alpha_0 \frac{2^{1-v}}{\Gamma(v)}
(\sqrt{2v} d_{ij})^v K_v(\sqrt{2v} d_{ij})
\label{Matern}
\end{equation}
where the $\Gamma$ is the gamma function, and the $K_v$ is the modified Bessel function of the second kind. The $v$ controls the smoothness, as GP using Matern kernel is $\lceil v \rceil - 1$ times differentiable in the mean-square sense \cite{GP, Matern}.
The distance $d_{ij}$ between two data points does not have to be Euclidean distance. A more general form would be:
\begin{equation}
d_{ij}^2 = \Sigma_{d=1}^D \frac{(x_{d, i} - x_{d, j})^2}{l_d^2}
\label{distance_metric}
\end{equation}
Where this distance metric can be anisotropic, suggesting distance along some axis contributes more uncertainty than others. 
The $l_d$ in the denominator is a positive number called length scale that determines how slowly we lose uncertainty (bigger length scale means the function is flatter) along the $d^\text{th}$ axis.

\subsection{Reset-free Trial-and-Error Learning}
The work of this paper is based on the Reset-free Trial-and-Error \cite{RTE} (RTE) algorithm. 
The RTE uses GP to learn the transformation model online.
This is a very clever choice for two reasons.
First, the online learning requires very high data-efficiency, which happens to be one of the strengths of GP. 
Secondly, the true dynamics happens at low level, while the repertoire-based method learns the transformation model at high level. This will inevitably introduce errors, which can be well incorporated by GP as being a probabilistic model.
In RTE, the task space is just the displacement of the robot in x and y direction after executing each policy.
The prior mean function of RTE is chosen to be the outcomes of the policies evaluated in the simulator, which are the results collected during the MAP-Elites policy generation. 
Since the policy outcome has two dimensions (x and y), RTE uses a different GP for each dimension.
The kernel is chosen to be the RBF kernel with an isotropic distance metric. The input space is chosen to be the same as the task space, namely the x and y displacement evaluated in the simulation. 
During online adaptation, the RTE first predicts the outcome of each policy in the repertoire using Eq. (\ref{GP_posterior}), then it uses a Monte Carlo Tree Search \cite{MCTS} (MCTS) to plan its actions.
It is worth noting that the RTE uses a periodic set of commands as the elementary policy. 
This means the policy execution is open-loop control, despite the robot knows its position and orientation.
While using such periodic controllers seems to be weaker than using a neural network, it is much more stable than using neural networks. 
This is because a periodic controller performs all its actions during evaluation, while neural networks are much more complicated and could behave very differently in different situations.
Since the repertoire based control aims to learn the distortion with limited number of interactions, the distortion cannot be too large or too complicated. Hence the use of periodic controller is a much better choice.

This method is not very ideal for three reasons. 
First, the use of outcomes in default simulation setting as the prior mean is not good enough. Since some conditions like broken leg can lead to very large distortions. In this case, such prior would be misleading. 
Second, the use of the task space as the input space is not ideal. Two policies leading to very close displacements could follow very different trajectories, hence they may react very differently to the same new environment.
Also, the kernel in the RTE is manually chosen, which depends on experiences of the designer and could lead to suboptimal results.
Nevertheless, this method successfully enabled a damaged hexapod robot to recover 77.52\% of its capability \cite{RTE} and significantly outperforms the baseline result that does not use the GP as the transformation model.
Some other works like Adaptive Prior selection for Repertoire-based Online Learning \cite{APROL} (APROL) take very similar approach as RTE but use several repertoires generated in different environments to address potential large distortions. 
Since RTE is easier to implement, and any improvements on the RTE can be easily transferred to other other works, it is selected as the algorithm that our work is based on.

\subsection{Dirichlet Process Mixture Model}
Dirichlet process mixture model (DPMM), also called infinite mixture model, is a Bayesian non-parametric model widely used in data clustering. 
It is based on a stochastic process called Dirichlet Process (DP).
DP assumes that the data are sampled from a distribution of distributions. For example, the data points come from a set of Gaussian distributions, while the mean and variance of each Gaussian follow another distribution. 
To make clear explanation, we first investigate the finite mixture model using DP, and then extend the derivations to the infinite limit.
We know that finite mixture models can be represented as:
p(\bm{y}) = \Sigma_{j=1}^k \pi_j \cdot p(\bm{y}|\bm{\theta_j})
\label{finite_mixture_model}
where this model consists of $k$ components, and $\bm{\theta_j}$ and $\pi_j$ are the parameters and the mixture weights (also called mixing proportions) for each mixture component.
DP assumes the parameters and the mixture weights are independently sampled. 
The parameters of the mixture components are sampled from a base distribution $H$, and the mixture weights are sampled from a symmetric Dirichlet distribution \cite{DP}:
p(\pi_1, \dots, \pi_k | \alpha) = 
\frac{\Gamma(\alpha)}{\Gamma(\alpha / k)^k} \prod_{j=1}^k \pi_j^{a/k - 1}
\label{Dirichlet_distribution}
Where the mixture weights are positive and sum up to 1. The constant $\alpha$ is called concentration parameter, which controls how dense the sampling will be.
To train a DPMM, the only thing we need to determine is the indicator of each data which points to the mixture component that this data belongs to. 
We then try to find the configuration of the indicators that maximizes the posterior likelihood (MAP estimate). 
This is very hard to do directly, typical alternative approaches include Gibbs sampling \cite{Gibbs_sampling, Gibbs} and the use of variational inference \cite{variational_method}.
This paper uses the Gibbs sampling method. 
In each iteration, we loop through each indicator and resample it based on the other indicators:
p(c_i = j|y_i, \bm{y}_{-i}, \bm{c}_{-i}, \alpha) 
\propto  \\
p(c_i = j|\bm{c}_{-i}, \alpha) \cdot
p(y_i|c_i=j, \bm{y}_{-i}, \bm{c}_{-i})
\label{indicator_posterior}
The subscript $-i$ means all the data except for $i$.
We know the probability of getting a certain configuration is:
p(c_1,\dots, c_n|\pi_1, \dots, \pi_k) = \prod_{j=1}^k \pi_j^{n_j}
\label{configuration_probability}
Where the $n_j$ in the superscript denotes the number of data assigned to the $j^\text{th}$ component.
Using Eq. (\ref{Dirichlet_distribution}) and standard Dirichlet integral, we can calculate the probability density of such configuration in the prior:
p(c_1, \dots, c_n | \alpha) =  \frac{\Gamma(\alpha)}
{\Gamma(\alpha / k)^k} \int \prod_{j=1}^k \pi_j^{n_j + a/k - 1} d\bm{\pi}
= \frac{\Gamma(\alpha)}{\Gamma(\alpha + n)} 
\prod_{j=1}^k \frac{\Gamma(n_j + \alpha/k)}{\Gamma(\alpha/k)}
\label{indicator_prior}
Hence we can find the posterior using Bayes rule:
p(c_i =j | \bm{c}_{-i} , \alpha) = 
\frac{n_{-i, j} + \alpha /k}{\alpha + n - 1}
\label{indicator_posterior_2}
So far, we have been discussing the case of finite mixture models. If we let the component number $k$ tend to infinity, we will have the equations for infinite mixture models:
p(c_i =j | \bm{c}_{-i} , \alpha) = 
\frac{n_{-i, j}}{\alpha + n - 1}
p(c_i \neq j \, \text{for any} \, n_j \neq 0 | \bm{c}_{-i} , \alpha) = 
\frac{\alpha}{\alpha + n - 1}
\label{indicator_posterior_3}
For any existing cluster (mixture component with at least one data assigned to it), this probability is proportional to the number of data in that cluster. 
Note that there is a non-zero probability that this data belongs to a new cluster. 
This is a very important property of DPMM that new clusters can be automatically generated based on the likelihood. 
To calculate the last term in Eq. (\ref{indicator_posterior}), we integrate the likelihood over the posterior:
p(y_i|c_i, \bm{y}_{-i}, \bm{c}_{-i}) = 
\int p(y_i|\bm{\theta})
p(\bm{\theta}|c_i, \bm{y}_{-i}, \bm{c}_{-i})
d\bm{\theta}
\text{where \,\,} 
p(\bm{\theta}|c_i, \bm{y}_{-i}, \bm{c}_{-i}) \propto
\,\, p(\bm{\theta})
\prod_{c_k=c_i, k \neq i} p(y_k|\bm{\theta})
\label{integral_1}
In the case of a new cluster, the integration is made over the prior:
p(y_i|c_i \neq j \, \text{for any} \, n_j \neq 0) = 
\int p(y_i|\bm{\theta})
p(\bm{\theta})d\bm{\theta}
\label{integral_2}
To conduct Gibbs sampling, we first start from an initial configuration of indicators (common choice is that every data is a cluster on its own).
Then in each iteration, we loop through each indicator and use Eq. (\ref{indicator_posterior}) to resample it. 
The $-i$ subscript on $\bm{y}$ means we need to remove the data from its current cluster during resampling, which will affect the result of Eq. (\ref{integral_1}) when calculating the probability of the data remaining in its current cluster. 
The Gibbs sampling is a Markov Chain Monte Carlo (MCMC) that satisfies irreducibility, positive recurrence and aperiodicity, hence it will eventually converge to the equilibrium distribution regardless of its initial state \cite{MCMC}. 
Since we are aiming to maximize the posterior likelihood, we need to calculate the likelihood after each iteration and record the configuration with the highest value.
The posterior likelihood is given by:
p(\bm{c}|\bm{y}) \propto 
\left[\prod_{n_j \neq 0} \alpha \Gamma(n_j)\right]
\prod_{i=1}^n p(y_i|c_i, \bm{y}, \bm{c}_{-i})
\label{posterior_likelihood}
Note that the probability in the second product doesn't have the $-i$ subscript on $\bm{y}$ like in Eq. (\ref{integral_1}), so the data doesn't need to be removed from its current cluster now.

We also need to define the base distribution $H$ that gives the prior distribution $p(\bm{\theta})$ and the concentration parameter $\alpha$ which controls the generation of new clusters.
Note that under the assumptions of DP, the final data distribution is a biased sample from $H$ unless $\alpha$ tends to infinity. Hence we don't need to ensure our $H$ being close to data distribution, while a rather conservative distribution is encouraged.
The $\alpha$ is selected based on Eq. (\ref{indicator_posterior_3}).
We can see that the probability of getting a new cluster is proportional to $\alpha$ and decreases asymptotically to zero as the number of data increases.
This is reasonable as the more data we have, the more certain we are that we have sampled at least one data from each cluster, hence the lesser we need a new cluster. 
If the $\alpha$ is infinite, DP believes all mixture component has the equal weight, and the data should always belong to a new situation regardless how many data we have collected.
In practice, we can estimate the amount of data we need in order to fully cover all situations. Then we can find the corresponding $\alpha$ using Eq. (\ref{indicator_posterior_3}) so that the probability decreases to a small value (e.g. 1\%) after we have collected this number of data.

DPMM is a very powerful tool to make non-parametric clustering without specifying the number of clusters. 
It can also easily incorporate the infinite limit than approaches working with finite models of unknown sizes like \cite{Bayesian_mixture}.
Comparing with optimization based methods like EM \cite{EM}, the use of MCMC can easily overcome the local optimal \cite{infinite_GMM}. 
For example, if there are duplicated mixture components, EM will still converge but the DPMM will eventually merge them as the Eq. (\ref{posterior_likelihood}) favours the data to be concentrated. 
Note that the above derivations of DPMM did not put any constrain on the type of distribution, namely the term $p(y|\bm{\theta})$ in Eq. (\ref{integral_1}). Such distribution can also be non-parametric distributions like Gaussian Processes. 
This means we can use DP to cluster the dynamics in RTE, which is GP, to build a mixture model of the real-world dynamics.

@@

%To formulate the problem clearly, we are aiming to leverage the collaboratively collected real-world interactions to assist a robot deployed in an unknown environment to quickly adapt and complete its task.
%We are provided with some episodes (each task-solving process corresponds to an episode of interaction data) of real-world interactions collected by several robots deployed in an unknown distribution of environments.
%The most intuitive idea is to label these historical data with their sources (like collected from grass terrain, collected with left leg broken, etc.) and build a simple database in the cloud server. 
%During the adaptation, the robot examines the current environment and uploads the result to the database server. The server then identifies the most similar environment recorded in the database and let the robot download the corresponding data to assist the its adaptation.
%However, accurately measuring and describing the environment can be expensive and difficult. It is also hard to determine whether a previously encountered environment shares similar dynamics with the current one even if we have accurate descriptions for them.
%In the case we do know that some historical data come from exactly the same environment as the current one, these data might be insufficient in quantity to assist the current task.
%Hence, it is reasonable to group the episodes resulting from similar dynamics into clusters, ensuring having enough data for analysis.
%We will assume that we don’t have access to any information of the environment and aim to conduct clustering only according to the dynamics.

@@

This is hard to do since we have little knowledge.
Even if we could give a perfect prior distribution, the integrals in Eq. (\ref{integral_1}) and Eq. (\ref{integral_2}) are intractable.
The solution is use Monte Carlo sampling in simulation combined with a trick called posterior convergence estimate.
We can see that these parameters are all related to the environment.
The Monte Carlo sampling is to collect a sample of these environment related parameters from the simulation as our prior distribution $H$.
To do this, we first randomize the simulation settings to get a sample of simulated environments.
We then collect a sufficient amount (e.g. 64) of interaction data from each environment and find the detailed parameters using maximum likelihood estimate (MLE):
\begin{equation}
\bm{\theta}^* = 
\arg \max_{\bm{\theta}} \, p(\bm{y}|\bm{\theta})
= \arg \min_{\bm{\theta}} \, -\log p(\bm{y}|\bm{\theta})
\label{MLE}
\end{equation}
The Monte Carlo sampling seems computationally expensive. But note that the sample amount for each environment is small and invariant to the size of the repertoire.
Hence it is much cheaper to do than finding the basis for the prior mean function, where we need to evaluate all the policies in several environments.

@@

\section{Experiment}

\subsection{Robot Setup}
To evaluate our methodology, experiment is conducted on the Ant robot in PyBullet simulation \cite{PyBullet} (see Fig \ref{Ant_robot}). 
This robot has 8 degrees of freedom (2 for each leg) corresponding to 8 motors powering its motion.
Note that the joints connecting the body and each leg can only rotate horizontally, meaning that the four thighs (the part of each leg directly attached to the body) are always on the same plane.
Hence, the Ant robot cannot lift up its leg to walk like real animals.
The motors employ PID positional control, where the torque is given by:
\begin{equation}
\tau = K_p \theta_e + K_i \int \theta_e(t) dt + K_d \frac{d \theta_e}{dt}
\label{PID}
\end{equation}
where $\theta_e = \theta_t - \theta$ ($\theta_t$ is the positional target).
In PyBullet, the maximum torque is $\pm 1$. 
For the motors on the body, $K_p$ is set to be 1.43 and $K_d$ is set to be 0.072.
For the motors on the knees, $K_p$ is 1.637 and $K_d$ is 0.082.
For all motors, $K_i$ is set to zero for simplicity. 
\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{intact_robot.pdf}
\caption{The Ant robot in PyBullet simulation. 
The four motors mounted on the body can rotate from $-40^{\circ}$ to $40^{\circ}$. The knee flexion angle for each leg ranges from 
$30^{\circ}$ to $100^{\circ}$.
}
\label{Ant_robot}
\end{figure}

Like in RTE, we use a periodic controller for the robot.
The period is set to 1 second, which corresponds to 100 time-steps in simulation.
The targets within each period is parametrized in the same way as in \cite{cully2015robots}.
The target of each motor is a smoothed squared wave (see Fig \ref{targets}) governed by three parameters: amplitude $\alpha$, phase $\phi$ and duty cycle $\tau$.
All the three parameters range from 0 to 1.
The duty cycle stands for the proportion of time in one period that the squared wave is in its high state.
This squared wave is then smoothed by a Gaussian kernel to remove rapid changes.
In each time-step of the simulation, we give each motor the corresponding positional target, which will then be converted into torques via PID.
Thus, each controller is parametrized by a vector of 24 dimensions, which is hence defined to be the genotype of this controller.
%
%
\begin{figure}[H]
\centering
\includegraphics[width=0.45\textwidth]{targets.pdf}
\caption{The periodic motor targets corresponding to $\alpha = 0.8$, $\phi=100$ ms and $\tau=0.6$.
The squared wave is smoothed by a Gaussian kernel with $\sigma = 100$ ms.
}
\label{targets}
\end{figure}

In this paper, we consider three sources of distortions: friction, payload and damages to the legs of the robot.
To collect the distortions as the basis for the prior mean function, we need to model these sources in simulation.
Different frictions can be easily achieved by changing the simulation settings.
Leg damages are modelled by giving a fixed positional target to the motor mounted on the knee.
This fixed target overwrites the command of the controller, so that the damaged leg is always in the upper position (see Fig \ref{damaged_robot}).
However, for unknown reasons, changing the mass of the robot do not affect the dynamics in simulation. 
Hence, we take an alternative approach by discounting the torque to model the payload.
%
%
\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{damaged_robot.pdf}
\caption{The robot walking with one leg (left blue-yellow leg) damaged.
The knee flexion angle of this leg is fixed at $30^{\circ}$ to reduce contact with the ground.
}
\label{damaged_robot}
\end{figure}

\subsection{Generating the Repertoire with MAP-Elites}
To implement MAP-Elites, the task space for our robot is defined to be its 2D displacement, spanning from -4.8 to 4.8 on both the $x$ and $y$ axes. 
This continuous space is then discretized into a $32 \times 32$ grid, yielding a resolution of 0.3 unit per cell.
To evaluate each genotype, we first decode it into the periodic targets and then execute them for 4 seconds.
The fitness is defined to be a reward of the lifespan (number of time-steps before the body of the robot hits the ground) penalized by the electrical cost \cite{QDgym}.
In each iteration of MAP-Elites, we generate 128 off-springs from 8 selected parents.

To design a crossover operation that balances exploitation and exploration,

The crossover operation should balance the exploitation and exploration by ensuring some inheritance of the parents to preserve good genes as well as some random variations for promoting diversity.

computation 10 cpu cores.

For each controller, apart from the 2D displacement, we also need to record the three parameters for arc-based dynamics representation calculated using Eq. (\ref{all_three_parameters}).

\subsection{Determine the Basis with SVD}

\subsection{Collecting Data Collaboratively}

\section{Results and Analysis}

\subsection{Dirichlet Process Clustering}

\subsection{Dynamics Learning}

\subsection{Tunnel Navigation Challenge}

\section{Conclusion}

%\begin{footnotesize}

\bibliographystyle{unsrt}
\bibliography{reference}
%\end{footnotesize}

%\end{multicols}
\end{document}

>>>>

@@

@@

This is hard to do since we have little knowledge.
Even if we could give a perfect prior distribution, the integrals in Eq. (\ref{integral_1}) and Eq. (\ref{integral_2}) are intractable.
The solution is use Monte Carlo sampling in simulation combined with a trick called posterior convergence estimate.
We can see that these parameters are all related to the environment.
The Monte Carlo sampling is to collect a sample of these environment-related parameters from the simulation as our prior distribution $H$.
To do this, we first randomize the simulation settings to get a sample of simulated environments.
We then collect a sufficient amount (e.g. 64) of interaction data from each environment and find the detailed parameters using maximum likelihood estimate (MLE):
\begin{equation}
\bm{\theta}^* = 
\arg \max_{\bm{\theta}} \, p(\bm{y}|\bm{\theta})
= \arg \min_{\bm{\theta}} \, -\log p(\bm{y}|\bm{\theta})
\label{MLE}
\end{equation}
The Monte Carlo sampling seems computationally expensive. But note that the sample amount for each environment is small and invariant to the size of the repertoire.
Hence it is much cheaper to do than finding the basis for the prior mean function, where we need to evaluate all the policies in several environments.

@@
